---
title: "STAT 545 - Assignment 1.3"
author: "Sean La"
date: "09/09/2019"
output:
  ioslides_presentation
---
## Loading the dataset

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

First, we'll load the Gapminder datatset.

```{r load_gapminder}
library(gapminder)
gapminder.df <- gapminder
```

## Simple summary statistics for life expectancy
Let's look at the structure of the Gapminder dataset...

\tiny
```{r gapminder_structure, size="huge"}
names(gapminder.df)
```
\normalsize
There's a column called `lifeExp`, which stands for life expectancy. 

---

Let's find out what the mean and standard deviations for life expectancy are!

```{r lifeExp_mean_sd}
print(paste('The mean life expectancy is', mean(gapminder.df$lifeExp), '.'))
print(paste('The standard deviation for life expectancy is', sd(gapminder.df$lifeExp), '.'))
```

## Histogram of life expectancies
Let's make a histogram of the life expectancies to visualize the data better.

```{r lifeExp_hist}
hist(gapminder.df$lifeExp)
```