---
title: "STAT 545 - Assignment 1.3"
author: "Sean La"
date: "09/09/2019"
output:
  ioslides_presentation
---
## Loading the dataset

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

First, we'll load the Gapminder datatset.

```{r load_gapminder}
library(gapminder)
gapminder.df <- gapminder
```

## Structure of data
Let's look at some of the data columns...

\tiny
```{r gapminder_structure}
names(gapminder.df)[1:4]
```
\normalsize
There's a column called `lifeExp`, which stands for life expectancy. 

## Summary statistics of life expectancy

Let's find out what the mean and standard deviations for life expectancy are!

```{r lifeExp_mean_sd}
life_exp_mean <- mean(gapminder.df$lifeExp)
print(paste('Mean life expectancy is',life_exp_mean,'.'))
life_exp_sd <- sd(gapminder.df$lifeExp)
print(paste('Standard deviation for life expectancy is',
            life_exp_sd,'.'))
```

## Histogram of life expectancies
Let's make a histogram of the life expectancies to visualize the data better.

```{r lifeExp_hist}
hist(gapminder.df$lifeExp)
```